\documentclass[11pt]{article}
\usepackage{comment}
\usepackage{graphicx}
\usepackage{fourier-orns}
\usepackage{algorithm, algpseudocode} % For pseudo code
\usepackage{tabto}
\usepackage{amsmath, amsthm, amssymb, amsfonts} % Linear algebra and logic related symbols
\usepackage{color}
\usepackage{fancyhdr} % Header
\usepackage{enumitem} % More option for enumeration and itemization
\usepackage{latexsym}
\usepackage{multicol}
\usepackage{fourier-orns}
\usepackage{algorithm, algpseudocode} % For pseudo code
\usepackage{tabto}
\usepackage{amsmath, amsthm, amssymb, amsfonts}
\usepackage{color}
\usepackage{fancyhdr} % Header
\usepackage{enumitem} % More option for enumeration and itemization
\usepackage{tikz}
\usetikzlibrary{arrows} % graph
\usepackage{url}
\topmargin=-.01in    %0cm
\textheight=9.0in     %24.1cm            
\evensidemargin=0in %0cm
\oddsidemargin=0.0in  %-.3cm
\textwidth=6.5in    %16.5cm 

\usepackage{amsmath}
\usepackage[skip=1pt]{caption}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{tabularx}

\title{MLH PE Fellowship Notes}
\author{Maninder (Kaurman) Kaur}

\begin{document}

\maketitle

\section*{Week 2: Linux Fundementals}

Linux is an OS specifically better for open applications, more versatile, and overall better privacy options. All terminal functions run on Linux commands, so let's learn that. \\
\noindent Scenario: System is Slow. To see why it's slow, they could use the \texttt{top} command to show summary info of list of processes or threads running on the Kernel. \\
Next, check the Disk IO using \texttt{iostat} that checks the input/output to see CPU utilization, number of transfer/sec, bulk reads/writes and CPU metrics. \\
Then, maybe they use \texttt{sar} that sees overall system performance. We could monitor things like CPU usage, Memory utilization, I/O device consumption, battery performance, etc. \\

\begin{table}[ht]
    \centering
    \caption*{Folders in Linux}
    \begin{tabular}{|c|l|}
        \hline
        / & primary directory \\
        /bin & essential executable programs \\
        /boot & files needed to boot system, such as kernal, boot config files, etc. \\
        /dev &  devices nodes, used to interact with hardware\\
        /etc & system-wide config files \\
        /home & user home directories \\
        /lib &  libraries required for /bin and /sbin \\
        /lib64 & 64-bit libraries \\
        /media & mount points for removable CDs, DVDs, etc. \\
        /mnt & temporarily mounted filesystems \\
        /opt & optional application software packages \\
        /proc & virtual pseudo-filesystem that gives info about system and processes \\
        /run & run-time variable data \\
        /sys & virtial pseudo-filesystem that gives info about system and processes \\
        /root & home directory of user \\
        /sbin & essential system binaries \\
        /srv & site-specific data \\
        /tmp & temporary files \\
        /usr & multi-user applications \\
        /var & variable data that changes system operation \\
        \hline
    \end{tabular}
\end{table}

\newpage

\noindent VPS (Virtual Private Server) that you can SSH into and work on. Reliable performance, data protection, control, etc.

\noindent Command line basics! This is where you tell the OS what to do. Remember, if you forget how a function works just do \texttt{man \{function name\}} or \texttt{--help}. \\

\noindent Moving: \texttt{cd} allows you to move between directories, \texttt{pwd} allows you to see your location, and \texttt{ls} lets you see the contents in the folder you're in. \\

\noindent Manipulating Files: \texttt{touch \{file name\}} lets you make a file, while \texttt{nano \{file name\}} lets you make a file and opens up and editor. \texttt{mkdir \{new folder name\}} makes a new directory while \texttt{rmdir} removes it. \texttt{cp} copies files, \texttt{mv} moves them, and \texttt{rm} removes them. \\

\noindent Viewing: \texttt{cat \{file name\}} prints all the contents in the file. \texttt{head} shows you the default top 10 lines, and you can add \texttt{head -n \{number\}} to see more or less specific. \texttt{tail} works the same, but shows you the ending. \\

\noindent Text Processing: \texttt{echo "message" > \{file name\}} allows you to print into something or simply onto the screen. \texttt{grep "find message" \{file name\}} returns every line that includes the message. \texttt{sed 's/\{text to replace\}/\{replacement text\}' \{file name\}} will replace all things from one message with another given a file name. To save the changes, add the tag \texttt{-i}. \texttt{sort \{file name\}} lexicographically sorts the file name from smallest to largest, and the extra argument after \texttt{| uniq} only shows distinct numbers. \\ \\

\noindent Boot Sequence steps:
\begin{itemize}
    \item BIOS(Basic I/O System)/UEFI locates and executes the boot program or boot loader
    \item Boot loader loads the kernal
    \item Kernal starts the init process (pid=1)
    \item Init manages system initialization, use systemd
\end{itemize}

\noindent BIOS contains all the code required to gain initial access to the keyboard, display screen, disk-drives, serial communications, etc. Typically placed in ROM chip (called ROM BIOS). 

\noindent Almost all Linux distros use GRUB (GRand Unified Bootloader). GRUB can boot multiple OSs, network-based diskless booting, etc.

\noindent System-wide config files are generally placed in \texttt{/etc} and their subdirectories, while ultra-specific are in home. Only text files should be in \texttt{/etc}. Files in \texttt{/etc/sysconfig} and it's subdirectories are used by many system utilities services. 

\noindent \texttt{shutdown} is used to bring the system down in a secure fashion.




\section*{Week 3: Scripting + Bash}
Scripting is designed to automate tasks in the Linux interface. A Shell script is essentially a text file that contains a sequence of commands for a UNIX based OS. A Python script is a plain text file with the python extension containing Python text to execute with an interpreter. 

\noindent A shell is a program that takes commands from user and gives it to the OS to process, an example of CLI (Command-Line Interface). Examples include, \texttt{bash, zsh, ksh, tcsh}. A terminal is a program that allows you to interact with the shell.

\noindent This is what the \texttt{date} command does.
\begin{lstlisting}[language=python]
$ date
Thu Jul 11 14:42:19 EST 2026
\end{lstlisting}

You will always need to start your bash script, ending in extension \texttt{.sh}, with this:
\begin{lstlisting}[language=python]
#!/bin/bash 
\end{lstlisting}

Here's an example where we can store a variable and also use the \texttt{echo} command:
\begin{lstlisting}[language=python]
#!/bin/bash

echo "What is your name?"
read PERSON
echo "Hello, $PERSON"

# in terminal
$ ./test.sh
What is your name?
Creed!
Hello, Creed!
\end{lstlisting}

\noindent To actually make the bash script executable, we use the following command: \texttt{chmod +x test.sh}.


\begin{table}[htbp]
    \centering
    \caption*{Relational Operators}
    \begin{tabularx}{\textwidth}{|c|X|X|}
        \hline
        -eq or = (string operators) & checks if value of two operands is equal or not & [\$a -eq \$b] is not true \\ \hline
        -ne or != (string operators) & checks if value of two operands are not true & [\$a -ne \$b] is true \\ \hline
        -gt & checks if the value of the left operand is greater than the value of the right operand & [\$a -gt \$b] is not true \\ \hline
        -lt & checks is the value of the left operand is less than the value of the right operand & [\$a -lt \$b] is true \\ \hline
        -ge & checks if the value of the left operand is greater than or equal to the value of the right operand & [\$a -ge \$b] is not true \\ \hline
        -le & checks if the value of the left operand is less than or equal to the value of the right operand & [\$a -le \$b] is true \\ \hline
        ! & logical negation & [!false] is true \\ \hline
        -o & logical OR & [1 -o 0] is true \\ \hline
        -a & logical AND & [1 -a 0] is false \\ \hline
        -z & checks if given string operand size is zero & [-z \$a] is not true \\ \hline
        -n & checks if given string operand size is non-zero & [-n \$a] is not false \\ \hline
        str & checks if str is not the empty string & [\$a] is not false \\
        \hline
    \end{tabularx}
\end{table}



\begin{table}[htbp]
    \centering
    \caption*{Unix File Test Operators}
    \begin{tabularx}{\textwidth}{|l|X|l|}
        \hline
        -b file & Checks if the file is a block special file; if yes, then the condition becomes true. & [ -b \$file ] is false. \\ \hline
        -c file & Checks if file is a character special file; if yes, then the condition becomes true. & [ -c \$file ] is false. \\ \hline
        -d file & Checks if the file is a directory; if yes, then the condition becomes true. & [ -d \$file ] is not true. \\ \hline
        -f file & Checks if file is an ordinary file as opposed to a directory or special file; if yes, then the condition becomes true. & [ -f \$file ] is true. \\ \hline
        -g file & Checks if the file has its set group ID (SGID) bit set; if yes, then the condition becomes true. & [ -g \$file ] is false. \\ \hline
        -k file & Checks if the file has its sticky bit set; if yes, then the condition becomes true. & [ -k \$file ] is false. \\ \hline
        -p file & Checks if the file is a named pipe; if yes, then the condition becomes true. & [ -p \$file ] is false. \\ \hline
        -t file & Checks if the file descriptor is open and associated with a terminal; if yes, then the condition becomes true. & [ -t \$file ] is false. \\ \hline
        -u file & Checks if the file has its Set User ID (SUID) bit set; if yes, then the condition becomes true. & [ -u \$file ] is false. \\ \hline
        -r file & Checks if the file is readable; if yes, then the condition becomes true. & [ -r \$file ] is true. \\ \hline
        -w file & Checks if the file is writable; if yes, then the condition becomes true. & [ -w \$file ] is true. \\ \hline
        -x file & Checks if the file is executable; if yes, then the condition becomes true. & [ -x \$file ] is true. \\ \hline
        -s file & Checks if the file has size greater than 0; if yes, then the condition becomes true. & [ -s \$file ] is true. \\ \hline
        -e file & Checks if file exists; is true even if file is a directory but exists. & [ -e \$file ] is true. \\ 
        \hline
    \end{tabularx}
\end{table}





\section*{Week 4: Services and Databases}
2 main databases + use cases:

\begin{itemize}
\item
  Relational databases: data stored in multiple, related tables. Adhere
  to ACID (Atomicity, Consistency, Isolation, and Durability). Usually
  structure data with predefined, fixed relations EXAMPLE: Oracle,
  MySQL, PostgreSQL\\
\item
  NoSQL (Non-Relational): data does not conform to predefined schema.
  Store unstructured data, or semi-structured. EXAMPLE: MongoDB
\end{itemize}

4 types of NoSQL databases:

\begin{itemize}
\item
  Doc-oriented: main unit is a document which is JSON containing
  different fields. Good use case would be good for data analysis and
  looking through records\\
\item
  Columnar: main unit is a column of a table, retrieved column wise.
  Log-base files good use case.\\
\item
  Key-value: all data has a unique key identifier. Good for dealing with
  cache\\
\item
  Graph: main unit contains nodes that represent entities and edges that
  represent relationships. Good from graph like structures and social
  networks.
\end{itemize}

Database management system (DBMS)\\
SQL (Structured Query Language) used on Relational DBMS

Cardinality - distinctiveness of info values contained in a column. High
cardinality means column contains an outsized proportion of distinctive
values. Low cardinality means more repeats.\\
Mapping constraint is a data constraint that express the \# of entities
to which another entity can be related via a relationship set. For
binary relationship set R on an entity A and B, there are four possible
mapping cardinalities:

\begin{itemize}
\item
  One to one 1:1\\
\item
  One to many 1:M\\
\item
  Many to one M:1\\
\item
  Many to Many M:M
\end{itemize}

Service - group of processes (known as daemons) running continuously in
background.Most Linux distros uses same process manager: systemd

\texttt{Systemd} is a system + service manager for Linux, allows
parallel startup of system services at boot time, on-demand activation
of daemons, and dependency based on service control logic.
\texttt{.service} -\textgreater{} system services, \texttt{.target}
-\textgreater{} group of systemd units. Main features:

\begin{itemize}
\item
  Socket-based activation allows parallel services to start without
  losing messages when restarted.\\
\item
  Starts services in parallel as soon as all listening sockets are in
  place. Significantly reduces time required to boot system\\
\item
  Only stops running services on shutdown
\end{itemize}

\texttt{systemctl} - command used to control and manage \texttt{systemd}
services\\
Activate - \texttt{systemctl\ start\ foo}, deactivate -
\texttt{systemctl\ stop\ foo}, restart -
\texttt{systemctl\ restart\ foo}, view status -
\texttt{systemctl\ status\ foo}. More commands -\textgreater{}
\texttt{systemctl\ status\ foo}

HTTP (hypertext transfer protocol) - set of rules for sending and
receiving web pages\\
Web request - communicative message that goes between client/web browser
to servers.\\
HTTP Communication happens between client and server. Client makes HTTP
request and gets back a HTTP response.

HTTP request line consists of Method, URL and Version\\
METHODS:

\begin{itemize}
\item
  GET: used to seek info from specified source.\\
\item
  HEAD: only returns headers\\
\item
  POST: seeks to create and make a new object\\
\item
  PUT: replace fully with new section

  URL: location that HTTP is referring to\\
  VERSION: what version? HTTP/3? HTTP/2? HTTP/1.1?\\
  HTTP request is followed by HTTP Header Structure:
\item
  General: applied on request and response w/o affecting database body\\
\item
  Request: contains info about fetched\\
\item
  Response: contains location of source requested\\
\item
  Entity: info about the body of resources
\end{itemize}

HTTP response has initial status line, some header lines and an entity
body\\
Status Codes:

\begin{itemize}
\item
  1xx: informational\\
\item
  2xx: success (200 is request was successful)\\
\item
  3xx: redirection\\
\item
  4xx: client error (404 is file not found, 400 is bad request/not in
  format system can read)\\
\item
  5xx: server error (500 is server error, 505 is requested HTTP version
  not supported)
\end{itemize}

Command: \texttt{curl\ http://example.org\ –head\ -silent}: curl is name
of command, --head or -I tells curl to send an HTTP request with head
method. And -silent tells curl to not display progress.

Caching - supplements primary database by removing unnecessary pressure
on it. Different types of caching:

\begin{itemize}
\item
  Database Integrated: managed within database engine and has built in
  write through capabilities. The cons are size and capabilities.\\
\item
  Local: stored within application, removing network traffic. Con is
  that each app has its own node, thus it cannot shared with other local
  caches.\\
\item
  Remote: stored in dedicated servers by NoSQL.
\end{itemize}


\section*{Week 5: Testing}
We need to make sure our code is always working as expected. Testing is conducted to ensure systems and applications are robust to prevent issues. We also text deployment so they must make sure CI/CD(Continuous integration/delivery/deployement) pipelines are ensured.

\noindent \textbf{Types of Testing:}
\begin{itemize}
    \item White Box Testing: internal structure, design and coding of software are tested to verify flow of I/O and to improve design, usability, and security. Code is visible to testers.
    \item Black Box Testing: functionalities of software applications are tested without having knowledge of internal code structure, implementation details, and internal paths. 
    \item Gray Box Testing: tests software product or application with partial knowledge if the internal structure of application.
\end{itemize}

\noindent \textbf{Testing Levels:}
\begin{itemize}
    \item Unit Testing: tests individual components of the software. Makes sure all the individual parts work as intended. It is ideal to do this test prior to Integration Testing. Advantages include: quality of code improved, able to find bugs easier, changing code becomes easier, easier to understand project through documentation, debugging process simplified.
        \begin{lstlisting}[language=python, mathescape=true]
# if this is code to calculate valume of Cube:
def cuboud_volume(l):
    return (l*l*l)

# then let's write a Unit Test!
import unittest
class TestCuboid(unittest.TestCase):
    def test_volume(self):
        self.assertAlmostEqual(cuboid_volume(2),8)
        self.assertAlmostEqual(cuboid_volume(1),1)
        self.assertAlmostEqual(cuboid_volume(0),0)
        self.assertAlmostEqual(cuboid_volume(5.5),166.375)
        \end{lstlisting}
    \item Integration Testing: tests the combination of different units, their interactions, subsystem unity, and code compliance with the requirements. Ideal for after Unit Tests. Advantages include: simultaneous testing of different modules becomes convenient, lifesaver for a separated team, covers large amount of code. To implement it, use \texttt{pytest} library.
    \item System Testing: tests the system as a whole. All the modules/components are integrated in order to verify the system works as expected or not. Ideal to do before giving it off as a final product or to go back to the drawing board. Advantages include: tests end-to-end scenarios to test the system. Tests the same environments as the production environment which understands perspective and prevents issues.
\end{itemize}

\noindent Always remember to start testing early! You should test a webpage like so:
\begin{itemize}
    \item Functionality: Test all links, test forms, and cookies are working as expected and not broken.
    \item Usability: test the visability and accessibility of the webpage and make sure the user can see and interact with everything easily. Navigation, content, etc.
    \item Interface: test to make sure the application is sending requests correctly to the database and outputting it right. Make sure the web server is handling the in between correctly and make sure all web queries appear in database.
    \item Database: test the database. Any errors in executing the queries, data integrity for creating, updating and deleting data, check response time, and test retrieval. 
    \item Compatibility: make sure the webpage displays correctly across all devices and browsers.
    \item Performance: test response times and different connection speeds, load test the application and normal and peak loads, stress test to a breaking point and test a crash response. 
    \item Security: test unauthorized access requests, accessing restricted files, check session that are killed for prolonged user activity, and use SSL certificates, websites should restrict to encrypted SSL pages.
    \item Crowd: get a large number of people(crowd) to execute tests. 
\end{itemize}

\noindent There are also different types of tests to unsure all system, applications, environments and infrastructure are functioning as expected.

\begin{itemize}
    \item Canary Releases: reduces risk of introducing new software version by slowly rolling out the change to a small subset of users first. 
    \item Load Testing: simulates high demand on a software by testing to see max operating capacity of the app, whether current infrastructure is sufficient, is it sustainable, how many users can it support.
    \item Feature Flagging and Dark Launches: sees which features and functions can be revealed to the user or no. They leave the features disabled, and slowly enable them.
    \item Application Monitoring: actively monitoring application. RRUM(Real User Monitoring) where we track real world interaction. We can also use Synthetic monitoring where it reacts from simulated users. 
    \item Tracing: we see a single interaction go through multiple layers and see if it follows the correct path. 
\end{itemize}

\noindent Test Driven Development(TDD) focuses on creating unit test cases before developing the actual code. Once new code passes test, it is refactored to an acceptable standard. There are 3 phases in TDD:

\begin{itemize}
    \item Create precise tests that verify functionality
    \item Correct the code once if fails
    \item Refactor the code once it passes to reduce redundancies or to optimize the code.
\end{itemize}

\noindent There are many benefits of TDD including creates more optimized code, helps developer better analyze and understand client requirements and requests, makes testing new functionalities easier, test coverage is higher to reduce later failure, and enhances overall productivity.

\noindent When we run the following script:
\begin{lstlisting}[language=python, mathescape=true]
$\$$ ls -l a\_file
-rw-rw-r-- 1 coop a\_project 1601 Mar 9 15:04 a\_file
\end{lstlisting}

\noindent The part, \texttt{rw-rw-r-}, meaning owner permissions(read and write), group permissions (read and write), and other permissions (read only). The user is \texttt{coop} and the group is \texttt{a\_group}

\noindent There are several types of permissions that can be given: \texttt{r} is read access, \texttt{w} is write access, and \texttt{x} is execute access. If not given, the placement is replaced by a dash. Other specialized permissions exist. \texttt{setuid} allows an executable to run with owner privileges, while \texttt{setgid} runs on group privileges.


\noindent If we want to change file permissions, we use \texttt{chmod}. Here's an example where we allows owner and world to have execute privileges and remove the group write permissions:
\begin{lstlisting}[language=python, mathescape=true]
$\$$ ls -l a\_file
-rw-rw-r-- 1 coop coop 1601 Mar 9 15:04 a\_file
$\$$ chmod uo+x, g-w a\_file
$\$$ ls -l a\_file
-rwxr--r-x 1 coop coop 1601 Mar 9 15:04 a\_file
\end{lstlisting}

\noindent The \texttt{u} stands for user(owner), \texttt{o} stands for other(world), \texttt{g} stands for group. You can also use octal digits for shorthand chmodding, using 3 digits for each person to give permission (user, group, world).
\begin{itemize}
    \item 1 - if the execute permission is desired
    \item 2 - if the write permission is desired
    \item 4 - if the read permission is desired
    \item 5 - means read/execute
    \item 6 - means read/write
    \item 7 - means read/write/execute
\end{itemize}

\noindent Whenever we need to change file ownership, we use \texttt{sudo chown [new user] [file]} and to change group ownership we use \texttt{sudo chgrp [new group] [file]}. Only superuser can do this.

\noindent POSIX ACLs(Access Control Lists) extend the simple user, group, and world system. 

\noindent Groups are defined in \texttt{/etc/group} where every line looks like this: \texttt{groupname:password:GID:user1,user2...}

\begin{itemize}
    \item Groupname is the name of the group
    \item password is the password placeholder, only may be set if \texttt{/etc/gshadow} exists
    \item GID is the group identifier. Values 0-99 are for system groups. Values between 100-GID\_MIN are considered special. Values above GID\_MIN are for UPG(User Private Groups).
    \item Then finally a list of all the members of the group.
\end{itemize}

\noindent The following may allow you to modify group settings:
\begin{itemize}
    \item groupadd - add a new group
    \item groupmod - modify a group and add new users
    \item groupdel - remove a group
    \item usermod - manage a user's group memberships
\end{itemize}

\subsection*{HackerRank Questions}

\noindent \textbf{Reverse a Linked List:} 
\begin{lstlisting}[language=python, mathescape=true]
"""
Complete the 'reverse' function below.
The function is expected to return an INTEGER_SINGLY_LINKED_LIST.
The function accepts INTEGER_SINGLY_LINKED_LIST llist as parameter.


For your reference:

SinglyLinkedListNode:
     int data
     SinglyLinkedListNode next
"""


def reverse(llist):
    # Write your code here
    prev = None
    current = llist
    
    while current is not None:
        next_node = current.next
        current.next = prev
        prev = current
        current = next_node
    
    return prev
    
    
if __name__ == '__main__':
    fptr = open(os.environ['OUTPUT_PATH'], 'w')

    tests = int(input())

    for tests_itr in range(tests):
        llist_count = int(input())

        llist = SinglyLinkedList()

        for _ in range(llist_count):
            llist_item = int(input())
            llist.insert_node(llist_item)

        llist1 = reverse(llist.head)

        print_singly_linked_list(llist1, ' ', fptr)
        fptr.write('\n')

    fptr.close()
    
\end{lstlisting}

\noindent \textbf{Cycle Detection:}
\begin{lstlisting}[language=python, mathescape=true]
"""
Complete the has_cycle function below.
For your reference:

SinglyLinkedListNode:
    int data
    SinglyLinkedListNode next
"""

def has_cycle(head):
    if head is None:
        return False
    
    slow = head
    fast = head
    
    while fast is not None and fast.next is not None:
        slow = slow.next
        fast = fast.next.next
        
        if slow == fast:
            return True
    
    return False

if __name__ == '__main__'
\end{lstlisting}

\end{document}
